Making a package from base R files

### 2022-01-19 PkgFromRbase/PkgFromRbase.Rmd
edits from AB

This article tries to explain an approach to developing alternative versions of
functions which are in the distributed base of R. Our interest was in developing
improvements to the nls() function and related features in R as part of a Google
Summer of Code project for which Arkajyoti Bhattacharjee was the funded student.
However, nls() has many tentacles involving a number of files and functions that
may or may not be called as `nls()` is executed.

@@

- build or install the two versions of R in separate directory trees and
  invoke them with separate launch scripts or programs, such as described
  in https://support.rstudio.com/hc/en-us/articles/215488098-Installing-multiple-versions-of-R-on-Linux ;

@@

```
nlspkg::Anyfn() # existing functionality or code
# and
nlsalt::Anyfn() # New functionality of code
```

@@

After we downloaded the ZIP file of Duncan's package from his Github repository, 
we built and checked the package. There were two WARNING messages:

@@

For Rstudio users, note that each package directory needs its own (name).Rproj
file, created using Rstudio's "File/New Project" etc. dialog. However, the package
directories with their Rproj files can be within the umbrella version control 
project `improvenls` for the Git version control. (We used a Gitlab
repository.)

@@

While all the above worked well in a Linux environment, when we tried running
the build and CHECK in a Windows environment, we found errors. (JN normally
runs in Linux Mint 20.x, but AB uses Windows 10. Each of us
has a VirtualBox VM for the alternative OS to allow for checking of our work.)
There is a load error for functions from the BLAS (Basic Linear Algebra Subroutines)
collection in a 32 bit Windows architecture. This is revealed by the Rstudio "Build / Clean and rebuild"
feature. Using the `CMD.exe` terminal window showed that `R CMD build` worked, but
that `R CMD INSTALL` gave the load errors.

@@

\url{https://stackoverflow.com/questions/42118561/error-in-r-cmd-shlib-compiling-c-code}

This suggested that a file `Makevars.win` is needed in the `src/` directory of the package(s),
and that this file should contain just one line

@@

Our packaged version of nls() from July 2021 is at 
https://gitlab.com/nashjc/improvenls/-/tree/master/nlspkg 
Note that this is drawn from the version of R of that date (R 4.1.0), so will not necessarily reflect
the distributed version of R when this document is read.

>>>>

This article tries to explain an approach to developing alternative versions of
functions which are in the distributed base of R. Our interest was in developing
improvements to the `nls()` function and related features in R as part of a Google
Summer of Code project for which Arkajyoti Bhattacharjee was the funded student.
However, `nls()` has many tentacles involving a number of files and functions that
may or may not be called as `nls()` is executed.

@@

- build or install the two versions of R in separate directory trees and
  invoke them with separate launch scripts or programs, such as described
  in \url{https://support.rstudio.com/hc/en-us/articles/215488098-Installing-multiple-versions-of-R-on-Linux};

@@

```
nlspkg::Anyfn() # existing functionality or code
# and
nlsalt::Anyfn() # New functionality of code.
```

@@

After we downloaded the ZIP file of Duncan's package from his GitHub repository, 
we built and checked the package. There were two WARNING messages:

@@

For RStudio users, note that each package directory needs its own (name).Rproj
file, created using RStudio's "File/New Project" etc. dialog. However, the package
directories with their Rproj files can be within the umbrella version control 
project `improvenls` for the Git version control. (We used a GitLab
repository.)

@@

While all the above worked well in a Linux environment, when we tried running
the build and CHECK in a Windows environment, we found errors. (JN normally
runs in Linux Mint 20.x, but AB uses Windows 10. Each of us
<!-- Decided abbreviations are obvious, so not explained. -->
has a VirtualBox VM for the alternative OS to allow for checking of our work.)
There is a load error for functions from the BLAS (Basic Linear Algebra Subroutines)
collection in a 32 bit Windows architecture. This is revealed by the RStudio "Build / Clean and Rebuild"
feature. Using the `CMD.exe` terminal window showed that `R CMD build` worked, but
that `R CMD INSTALL` gave the load errors.

@@

\url{https://stackoverflow.com/questions/42118561/error-in-r-cmd-shlib-compiling-c-code}.

This suggested that a file `Makevars.win` is needed in the `src/` directory of the package(s),
and that this file should contain just the following line.
<!-- Changed sentence to avoid period in line below which might cause errors if left in code. -->

@@

Our packaged version of `nls()` from July 2021 is at

\url{https://gitlab.com/nashjc/improvenls/-/tree/master/nlspkg}.

Note that this is drawn from the version of R of that date (R 4.1.0), so will not necessarily reflect
the distributed version of R when this document is read.
